---
title: "Problem Set 7"
output:
html_document: default
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# <span style="color:blue"> *Exercise 7.2* </span>

## <span style="color:orange"> _part a)_ </span> 

</br>

#### <span style="color:green"> Creating the data frame </span>

```{r}
df <- data.frame("nearsightedness" = rep (c("yes","no"), times = c(21,173)))
```


</br>

## <span style="color:orange"> _part b)_ </span> 

</br>

#### <span style="color:green"> Show Table </span>

```{r}
table(df)
```

</br>

## <span style="color:orange"> _part c)_ </span> 

</br>

#### <span style="color:green"> Hypotheses </span>

H0: p = 0.08 (near-sightedness does affect about 8% of all children, the data does not provide enough evidence that the 8% value is inaccurate)
HA: p != 0.08 (near-sightedness does not affect about 8% of all children, the data does provide enough evidence that the 8% value is inaccurate)

</br>

## <span style="color:orange"> _part d)_ </span> 

</br>

#### <span style="color:green"> Load necessary packages </span>

```{r, message=FALSE}
library("infer")
library("tidyverse")
```


#### <span style="color:green"> Set seed for reproducable results & Create a null distribution </span>

```{r, message=FALSE}
set.seed(12345)

null_distribution <- df %>%
  
                        specify(response = nearsightedness, success = "yes") %>%
                          
                          # point estimate/ population parameter
                          hypothesize(null = "point", p = 0.08 )  %>% 
                          
                            generate(reps = 1000) %>% 
    
                              calculate(stat = "prop")
    
           
```

#### <span style="color:green"> View null distribution as tibble </span>

```{r}
tibble(null_distribution)
```

</br>

## <span style="color:orange"> _part e)_ </span> 

</br>


#### <span style="color:green"> Visualize the simulated null distribution </span>

```{r}
null_distribution %>%
                  visualise() + 
                    shade_p_value(obs_stat = 21/194, direction = "both")
```

#### <span style="color:green"> Corresponding p-value of the test </span>

```{r}
null_distribution %>% get_pvalue(obs_stat = 21/194, direction = "both")
```

</br>

## <span style="color:orange"> _part f)_ </span> 

</br>

#### <span style="color:green"> Conclusion </span>

Since the corresponding p-value of the test (0.198) is greater than the significance level (0.05) chosen before calculating the p-value, we do not have sufficient evidence to reject the H0 hypothesis which stated that near-sightedness does affect about 8% of all children.

